DATA 621 - Business Analytics and Data Mining

### 2020-11-10 HW4/HW4_Draft.Rmd
model 3 updates

```{r loadData, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(mice)
library(arm)
library(corrplot)
library(data.table)

@@

@@

@@

## Model 3
### Frequency Model
### Severity Model

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r zeroVal}
```
```{r statsN}
```
```{r graphsN}
```
```{r graphsF}
```
```{r corrgram}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r m2clean}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```
```{r model2RunFreq}
```
```{r model2RunSev}
```
```{r setUpCompTables}
```
```{r validateM1F}
```
```{r validateM1S}
```
```{r validateM2F}
```
```{r validateM2S}
```
```{r freqComp}
```
```{r sevComp}
```

>>>>

```{r loadData, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(mice)
library(arm)
library(corrplot)
library(data.table)
library(tidyverse)

@@

## Model Set #3
### Missing Data
Similar to model 2, model set 3 deals with missing numerical data by imputing 
them using the Multivariate imputation by chained equations (MICE) method. Multiple 
imputation involves creating multiple predictions for each missing value, helping 
to account for the uncertainty in the individual imputations.

One thing to note about Model 3 is that it does *not* assume that an `NA` in 
the `HOME_VAL` variable is indicative of renting. This is a key difference between
this model and the previous 2.

```{r m3clean}
# Model 3 - data transformation
# MICE imputation on train and test set
m3train <- complete(mice(data = trnX %>% select(-INDEX), method = "pmm", print=FALSE),3)
m3train$TARGET_FLAG <- factor(m3train$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))

m3test <- complete(mice(data = tstX %>% select(-INDEX), method = "pmm", print=FALSE),3)
m3test$TARGET_FLAG <- factor(m3test$TARGET_FLAG, levels = c(0, 1), 
                     labels = c("NoClaim", "Claim"))

```

@@

## Model Set #3
### Frequency Model
The frequency model is used to determine whether the vehicle is likely to crash.
The base of this is a binary classifier, with the output being a probability 
representing the likelihood of crashing.

#### Baseline Model
We will start with a simple logistic model to serve as a baseline. This includes
all variables in the dataset.

```{r model3RunBaseLog}
# Base log model 3
m3BaseLog <- glm(TARGET_FLAG~., family = binomial, data = m3train %>% select(-TARGET_AMT))
kable(summary(m3BaseLog)$coefficients, digits = 3L,
      caption = 'Model 3 - Base Logistic Regression Output')

```

We can immediately see that a few variables *exceed* the 0.05 p-value threshold
for significance.

#### Enhanced Frequency Model 
Backwards stepwise regression is performed and the result is a model with the 
following variables *removed*: `AGE`, `HOMEKIDS`, `YOJ`, `RED_CAR`, and `CAR_AGE`.

```{r model3RunFreq}
# Final log model 3
train_control <- trainControl(method = "cv", number = 10)
m3FreqFit <- train(TARGET_FLAG ~ KIDSDRIV+INCOME+PARENT1+HOME_VAL+MSTATUS+SEX+EDUCATION+JOB+
              TRAVTIME+CAR_USE+BLUEBOOK+TIF+CAR_TYPE+OLDCLAIM+CLM_FREQ+REVOKED+MVR_PTS+
              URBANICITY,
              data = m3train %>% select(-TARGET_AMT),
              trControl = train_control,
              method = "glm",
              family=binomial())
kable(summary(m3FreqFit)$coefficients, digits = 3L,
      caption = 'Model 3 - Final Logistic Regression Output')

```

#### Coefficient Discussion 
Some interesting observations from the final results:

* Living in an urban area has the greatest impact on an individual's 
likelihood of getting into a crash. This makes sense - urban areas are more 
crowded than rural areas, so the odds of crashing are higher. 
* If your car is used for commercial purposes you have a higher probability
of getting into a crash. 
* As expected, if you've had your license revoked within the past 7 years, 
you're more likely to get into a crash. 
* Being a female makes you less likely to get into a crash. 
* Having a sports car makes you more likely to get into a crash than having 
either an SUV, Van, Pickup, or Panel Truck. 
* Being a manager or doctor make you less likely to get into a car crash. 
Conversely, you're more likely to crash if you're a student, blue collar worker, 
clerk, homemaker, professional, or lawyer.

### Severity Model
The severity model is used to determine how much the crash will cost the
insurance company. Linear regression will be used to estimate the total value
of the cost for cars that are known to have been in an accident and accrued
damage costs.

#### Base Model
We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset, but only for the records that have
a TARGET_AMT > $0.

```{r model3RunBaseLin}

m3BaseLin <- lm(TARGET_AMT ~ ., data=m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))
knitr::kable(summary(m3BaseLin)$coefficients, digits = 3L,
             caption = 'Model 3 - Base Linear Regression Output')

```

A look at the final output shows us that only 1 variable is deemed
important (`BLUEBOOK`).

#### Enhanced Model
In this case, if we were to use backwards stepwise regression, we would
be left with a model that is determined by only 1 variable. However,
intuitively we know that some of the variables are likely to effect
the final payout. For this reason, we will choose to include: `BLUEBOOK`,
`CAR_AGE`, and `CAR_TYPE`.

```{r model3RunSev}

m3SevFit <- lm(TARGET_AMT~ BLUEBOOK+CAR_AGE+CAR_TYPE, data = m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))

knitr::kable(summary(m3SevFit)$coefficients, digits = 3L,
             caption = 'Model 3 - Final Linear Regression Output')

```

#### Coefficient Disucssion
The intercept of the final model tells us that the base cost of a crash is a 
little under $5000. Some results:

* The positive coefficient for the `BLUEBOOK` variable means that the higher
the bluebook value for the vehicle, the more the payout will be. This makes
sense. 
* The negative coefficient for the `CAR_AGE` variable means that the older
the vehicle is, the less the payout is. Once again, this makes sense. 
* Non-intuitively, the positive coefficient on the Van option for the 
`CAR_TYPE` variable means that van collisions have a higher payout. All 
other options for this variable (including a sportscar) result in a 
lower payout.

@@

## Model 3
### Frequency Model
```{r validateM3F}

# Model 3 Frequency: predict on validation data
m3Fpred <- predict(m3FreqFit, m3test %>% select(-TARGET_AMT))
m3FpredProb <- predict(m3FreqFit, m3test %>% select(-TARGET_AMT), type = 'prob')

# Model 3 Frequency: Calculate accuracy statistics
m3FCM <- confusionMatrix(m3Fpred, m3test$TARGET_FLAG, mode = 'everything')
freqCompTable[3, 2] <- m3FACC <- sum(diag(m3FCM$table)) / sum(m3FCM$table)
freqCompTable[3, 3] <- m3FF1 <- m3FCM$byClass[7]
freqCompTable[3, 4] <- m3FAUC <- auc(response = m3test$TARGET_FLAG,
                                     predictor = m3FpredProb[, 1])
m3FCM$table

```
Frequency Model 3 returns an accuracy of `r round(m3FACC, 4)`, an F1 score of
`r round(m3FF1, 4)`, and an AUC of `r round(m3FAUC, 4)`.

```{r validateM3S}
# Model 3 Severity: subset data
m3testFinal <- m3test %>% 
  select(-TARGET_FLAG) %>%
  filter(TARGET_AMT > 0)

# Model 3 Severity: predict on validation data
m3Spred <- predict(m3SevFit, newdata = m3testFinal)

# Model 3 Severity: Calculate accuracy statistics
m3Ssum <- postResample(m3Spred, tstSev)
sevCompTable[3, 2] <- m3SRMSE <- m3Ssum[1]
sevCompTable[3, 3] <- m3SMAE <- m3Ssum[3]
sevCompTable[3, 4] <- m3SR2 <- m3Ssum[2]
```

Severity Model 3 returns an RMSE of `r round(m3SRMSE, 1)`, an MAE of
`r round(m3SMAE, 1)`, and an \(R^2\) of `r round(m3SR2, 4)`.

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r zeroVal}
```
```{r statsN}
```
```{r graphsN}
```
```{r graphsF}
```
```{r corrgram}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r m2clean}
```
```{r m3clean}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```
```{r model2RunFreq}
```
```{r model2RunSev}
```
```{r model3RunBaseLog}
```
```{r model3RunFreq}
```
```{r model3RunBaseLin}
```
```{r model3RunSev}
```
```{r setUpCompTables}
```
```{r validateM1F}
```
```{r validateM1S}
```
```{r validateM2F}
```
```{r validateM2S}
```
```{r validateM3F}
```
```{r validateM3S}
```
```{r freqComp}
```
```{r sevComp}
```
